Freezing of a Weeks-Chandler-Anderson Fluid using classical Density Functional Theory

### 2020-11-27 papers/thesis-kirstie/thesis.tex
more edits

\documentclass[double,12pt]{beavtex}
\usepackage{graphicx}
\usepackage{rotating} %Package added to allow the rotation of figures and chart on a page, {sidewaysfigure} command
\usepackage{tablefootnote} %Packaged added to allow footnotes in the tabular environment, use \tablefootnote command
\title{Freezing of a Weeks-Chandler-Anderson Fluid using classical Density Functional Theory}
\author{Kirstie L. Finster}
\degree{Master of Science}
\doctype{Thesis}
\department{Physics}
\depttype{Department}
\depthead{Head}
\major{Physics}
\advisor{David Roundy}
\submitdate{October ??, 2020}
\commencementyear{2021}

@@

\date{October 2020}

\abstract{
  We use Classical Density Functional Theory (cDFT) to predict the 
  freezing of a Weeks-Chandler-Anderson (WCA) fluid. A WCA fluid 
  consists of soft spheres that exhibit the repulsive portion of the 
  Lennard-Jones potential widely used to model the interaction of two 
  neutral atoms. Our cDFT functional is constructed using Soft Fundamental 
  Theory (SFMT) with Tarazona's tensor expression for $\Phi_3$, and Schmidt's
  erf model in which we have incorporated temperature dependent
  parameters to fit the results to those predicted by Monte-Carlo simulations.
  Our functional was successful in predicting the freezing of the WCA fluid 
  with a fairly good match to Monte-Carlo simulations. This result suggests 
  that SFMT may provide a solid foundation for future density functionals 
  that will model real fluids.
}

@@

Soft Fundamental Measure Theory (SFMT) will be used to form the 
functional for the excess Helmholtz free energy. 
SFMT is a method that makes the task of 
forming an expression for the excess Helmholtz free energy of a particular 
model fluid much easier since the interaction that each sphere, or atom, 
has with every other sphere does not have to be taken explicitly into account.
Once the functional for the Helmholtz free energy is formed, classical 
Density Functional Theory is used to find the equilibrium 
Helmholtz free energy at a particular temperature and density. 
According to cDFT the equilibrium 
Helmholtz free energy of the system can be found by varying the number 
density profile $n(\vec{r})$ until the free energy is minimized. 
Results will be compared to those from Monte-Carlo simulations.

@@

The FMT equations developed so far have been for a Hard-sphere fluid.
All of the weight functions, the relations between them, and the equation 
for the derivative of the mayer function, can be expressed in terms of 
$w_{2}(r)$ which, for a Hard-sphere fluid, consists of a delta function.
The weight function $w_{2}(r)$ is in turn related to $w_{3}(r)$ through 
Equation~\ref{w2_w3_relation}.
Soft Fundamental Measure Theory extends FMT discussed so far to fluids 
with soft potentials by keeping the relationships between the weight 
functions the same, but replacing $w_{3}(r)$ with a weight function 
appropriate for a soft sphere potential~\cite{schmidt1999density}. 
This is what we do to form a functional for the Weeks-Chandler-Anderson 
fluid as detailed in the next chapter on Methods.

@@

Our goal is to generate temperature-density and pressure-temperature 
phase diagrams for a Weeks-Chandler-Anderson model fluid using 
classical Density Functional Theory. 
A temperature-density phase diagram for a fluid that does not
have attraction between the particles that make up the fluid will show
two distinct regions - one at lower densities, and one at higher densities,
separated by a coexistence region in which there is found a mixture of the 
fluid at some fixed lower density, and at some fixed higher density. An 
abrupt change from some maximum lower density to some minimum higher density 
occurs for certain combinations of pressure and temperature which form
a line on a pressure-temperature phase diagram. This line distinguishes 
the liquid region which exhibits lower densities from the 
solid which exhibits higher densities.

The transition between the liquid 
and solid state is abruptly inhomogeneous in terms of the number density, 
and we study the behavior 
of the fluid by treating the liquid as having a homogeneous number density, 
and by constructing a crystal lattice in real space for the solid. 
It is known that a fluid with a soft potential like the WCA fluid will 
form a Face Centered Cubic (FCC) crystal in its solid state~\cite{Hansen}. 
Although the entire 
crystal lattice can be constructed with FCC cells, 
FCC cells are not the smallest cells with which the lattice can be built. 
The shaded parallepiped shown within the FCC cell in Figure \ref{fig:primitivecell} 
is the smallest structure that, when repeated, re-creates the entire crystal 
lattice. This parallepiped primitive cell takes up one-fourth the volume 
of an FCC cell, and contains the equivalent of one atom.
It includes portions of eight atoms each located on a lattice site conveniently 
reachable by moving along one of three different lattice vectors to step 
from one atom to the next. The three lattice vectors will be useful in 
creating the lattice structure computationally, and will also make the 
program run faster.

  \begin{figure}[h!]
    \centering
    \includegraphics[height=3.5cm]{PrimitiveCellLightBlue.png}
    \caption{The Primitive Crystal Cell is shown by the shaded parallelpiped}
    \label{fig:primitivecell}
  \end{figure}

In a large collection of cells, the atom at each lattice site
will not realistically reside exactly at the lattice point. 
Statistically there will typically be a normal, or Gaussian distribution 
of positions about the lattice point as pictured in the far right box in 
Figure \ref{fig:Ensemble_Gaus}. 
A plot of a Gaussian distribution in 2-dimensions is shown in Figure~\ref{fig:Gaus_plot}
where the width of the Gaussian is described by $\sigma$.  
A 3-dimensional Gaussian distribution centered at a lattice point can be used to represent 
a number density profile $n(\vec{r})$  which describes how the atoms are 
statistically distributed over a collection of cells. 
It is important to note that the number density does not change when introducing a number 
density profile. There are still the same number of atoms over some fixed volume. 
\begin{equation}{n=<n(\vec{r})>=\frac{1}{V}\int{n(\vec{r})}{d\vec{r}}=\frac{N}{V}=\frac{1~\text{atom}}{V}}\end{equation}
where $n(\vec r)$ is itself an ensemble average over all microstates, and is independent of time.
\begin{equation}n(\vec r)=~<\sum_{i=1}^N\delta(\vec r - \vec r_i)>~\end{equation}

  \begin{figure}[h!]
    \centering
    \includegraphics[height=2.5cm]{Ensemble_Gaussian.png}
    \caption{A Gaussian Distribution Number Density Profile}
    \label{fig:Ensemble_Gaus}
  \end{figure}

@@

It may be that not all lattice sites are occupied. Vacancies can be taken 
into account while keeping the overall number density the same by 
making the cells smaller. A large collection of smaller cells creates a 
new number density profile shown in Figure~\ref{fig:Ensemble_Smallcells} 
with eight smaller Gaussians each of which is now one-eighth the height. 
Instead of thinking of an atom in one of the eight small cells, the same 
Gaussian distribution can be obtained from thinking of one-eighth of an 
atom in each small cell as shown in Figure~\ref{fig:SameStatPic}. 
This makes it convenient to describe the new 
number density in terms of the fraction of vacancies, $f_v$: 
\begin{displaymath}{ n = (1-f_v){n_{no.vacancies}}}\end{displaymath}

 \begin{figure}[h!]
    \centering
    \includegraphics[height=2.5cm]{Ensemble_Smallcells.png}
    \caption{A Gaussian Distribution Number Density Profile.}
    \label{fig:Ensemble_Smallcells}
  \end{figure}

@@

By varying the width of the Gaussians and the fraction of vacancies, various
number density profiles can be generated as shown in Figure~\ref{fig:Ensemble_vary}.
Once again, it is important to note that the number density is the same 
for all these number density profiles $n(\vec{r})$. 
A homogeneous number density profile is used for a liquid is shown in Figure \ref{fig:homogen_denisty}.

\begin{figure}[h!]
    \centering
    \includegraphics[height=8cm]{VaryWidthandVacancies.png}
    \caption{Several Gaussian Distribution Number Density Profiles all 
    with the same number density, that is, there is one atom over the fixed volume.}
    \label{fig:Ensemble_vary}
  \end{figure}

 \begin{figure}[h!]
    \centering
    \includegraphics[height=4cm]{figs/homogeneous_bold-box.pdf}
    \caption{A homogeneous number density profile $n(\vec{r})$.}
    \label{fig:homogen_denisty}
  \end{figure}

For a given combination of temperature and number density, a suite of number 
density profiles are generated similar to those represented in the example 
set of profiles shown in Figure~\ref{fig:Ensemble_vary}. 
The Helmholtz free energy $F[n(\vec{r})]$ is calculated for each number 
density profile, and also for the homogeneous number density profile 
characteristic of a liquid. Both the ideal and excess free energies are 
found, and added together to get the total Helmholtz free energy.
According to classical Density Functional Theory, of all the Helmholtz 
free energies calculated is the minimum Helmholtz free energy, and its 
corresponding number density profile that are the actual equilibrium values 
for that temperature and number density.
The process is repeated to find the free energies at other combinations 
of temperatures and number densities.

Once the Helmholtz free energy for both the crystal and liquid states of 
the fluid are found for a given temperature and density, they can be 
used to construct phase diagrams.
Outside the region of coexistence which exists between the liquid and 
crystal states, the fluid will be completely in a crystalline state if the 
crystal free energy is lower than the liquid free energy, and it will be 
completely in a liquid state if the liquid free energy is lower than 
the crystal free energy. Maxwell's Double Tangent Construction is used 
to determine the actual number densities of the crystal and liquid states
at the boundaries of the coexistence region for a given temperature, 
and construct phase diagrams as described in the section on The Theory of
Freezing and Maxwell's Double Tangent Construction.

@@

Phase diagrams for the WCA fluid are shown in Figures~\ref{fig:Phase_Diagram_of_T_vs_n} 
and \ref{fig:Phase_Diagram_P_vs_T}. In all cases $\sigma=1$ and $\epsilon=1$. 
In Figure~\ref{fig:Phase_Diagram_of_T_vs_n} the reduced temperature 
$T^*=\frac{kT}{\epsilon}$ is plotted against the reduced number density 
$n^*=n\sigma^3$ for temperatures of kT = 0.5 to 3.0. The red area is fluid, 
and the blue area is crystal. The gray area that separates the fluid, 
and the crystal regions is where some fluid and crystal exist together 
at varying proportions. The region of coexistence slants slightly to the 
right toward higher density as the temperature increases. 
This is expected since at higher temperatures, the fluid would have to 
be under greater compression before forming a solid. The white lines show 
results from Monte-Carlo simulations of the WCA fluid~\cite{May}. 
Overall, the general shape, and position of the lines correspond
fairly closely to the edges of the region of coexistance predicted from 
the DFT method, but the width of the coexistence region is about one-third 
larger for the Monte-Carlo results for most temperatures. The white line 
on the right, representing the density of the solid at each temperature, 
is in fairly good agreement with our DFT results for all temperatures. 
However, the white line on the left, representing the density of the 
liquid at each temperature, is consistently farther left than the liquid 
transition densities our method predicted.

%Figure~\ref{fig:Phase_Diagram_of_T_vs_n_AS} shows our results for a wider 
%range of reduced temperatures from kT=0.5 to 38. The white lines show the results
%of molecular simulations of the WCA fluid made by Ahmed and Sadus~\cite{ahmedsadus}.
%As can be seen in the figure, the region of coexistence in our simulations 
%falls within the region of coexistance predicted by Ahmed and Sadus at almost 
%all temperatures from approximately kT=0.5 to 28. At the low temperature of kT=0.5,
%the width of the coexistance regions are nearly the same, while moving to 
%higher temperatures,the width of our coexistance region ranges from about 
%one-half to one-forth the size. Although, our coexistence region is only 
%about one-forth the size at kT=28, it is centered within the coexistence 
%region predicted by Ahmed and Sadus.

\begin{figure}
  \centering
  \includegraphics[height=10cm]{figs/Phase_Diagram_of_T_vs_n}
  \caption{Phase Diagram of the reduced temperature verses the reduced density for the WCA fluid. 
  White lines show the liquid and solid transition densities at each temperature from 
  Monte-Carlo simulations of the WCA fluid. The gray region is the region of coexistence.}
  \label{fig:Phase_Diagram_of_T_vs_n}
\end{figure}

\begin{figure}
  \centering
  %\includegraphics[width=\columnwidth]{figs/Phase_Diagram_of_T_vs_n_AS}
  %\includegraphics[height=10cm]{figs/Phase_Diagram_of_T_vs_n_AS}
  %\caption{Phase Diagram of the reduced temperature verses the reduced density for the WCA fuild. 
  %White lines show the liquid and solid transition densities at each temperature from 
  %simulations of a WCA fluid by Ahmed and Sadus~\cite{ahmedsadus}. The gray region is the region of coexistence.}
  \label{fig:Phase_Diagram_of_T_vs_n_AS}
\end{figure}

Figure~\ref{fig:Phase_Diagram_P_vs_T} shows the reduced pressure $p^*=\frac{P\sigma^3}{\epsilon}$ 
plotted against the reduced temperature for temperatures of kT = 0.5 to 3.0. Again, the red area 
is fluid, and the blue area is crystal. The general trend of the values for the transition 
temperatures and pressures is what would be expected, with the WCA fluid becoming a solid at
a higher pressure for a higher temperature, or becoming a liquid at a higher temperature 
for a higher pressure. But the transition pressures and temperatures are consistently higher 
than the Monte-Carlo simulations shown by the white line, and the discrepancy grows as 
the temperature increases. 
%Figure~\ref{fig:Phase_Diagram_P_vs_T_AS} shows our results for a larger 
%range of reduced temperatures from kT=0.5 to 38, compated with the results
%from Ahmed and Sadus's molecular simulations. Here again, our transition 
%pressures and temperatures are consistently higher.

\begin{figure}
  \centering
  \includegraphics[height=10cm]{figs/Phase_Diagram_of_P_vs_T}
  \caption{Phase Diagram of the reduced pressure verses the reduced temperature. 
  The white line marks the boundary from Monte-Carlo simulations.}
  \label{fig:Phase_Diagram_P_vs_T}
\end{figure}

\begin{figure}
  \centering
  %\includegraphics[width=\columnwidth]{figs/Phase_Diagram_of_P_vs_T_AS}
  %\includegraphics[height=10cm]{figs/Phase_Diagram_of_P_vs_T_AS}
  %\caption{Phase Diagram of the reduced pressure verses the reduced temperature. 
  %The white line marks the boundary from simulations by Ahmed and Sadus~\cite{ahmedsadus}.}
  \label{fig:Phase_Diagram_P_vs_T_AS}
\end{figure}

%Figure~\ref{fig:p-vs-T_at_fixed_density} shows individual plots of the 
%reduced pressure versus the reduced temperature at fixed reduced densities 
%from $n^*$ = 0.7 to 1.2. As can be seen by comparison with Figure 
%\ref{fig:Phase_Diagram_P_vs_T_AS}, the plotted lines display kinks when 
%the WCA fluid transistions into a solid. These downward bends align to 
%form the curved interface separating the blue and red areas in 
%Figures~\ref{fig:Phase_Diagram_P_vs_T} and \ref{fig:Phase_Diagram_P_vs_T_AS}.

\begin{figure}
  \centering
  %\includegraphics[width=\columnwidth]{figs/p-vs-T_at_fixed_n_noMC}
  %\caption{Plot of the reduced pressure verses the reduced temperature. 
  %The lines bending downward indicate freezing.}
  \label{fig:p-vs-T_at_fixed_density}
\end{figure}

\chapter{Conclusion}
In this paper we were successful in predicting the freezing of a 
Weeks-Chandler-Anderson fluid using our SFMT functional built by 
forming temperature dependent parameters $\alpha(T)$ and $\Xi(T)$ to use 
in Schmidt's erf model to better fit our results to those expected for a
WCA fluid. We were able to generate a temperature-density 
phase diagram that yielded coexistence regions that approximately 
matched the Monte-Carlo
% and molecular simulation 
results. 
%for values of kT=0.5 to 3. 
%kT=0.5 to 38
We are also able to generate a pressure-temperature
phase diagram with a shape that is expected, but the transition 
temperatures and pressures were too high.

>>>>

\documentclass[double,12pt]{beavtex}
\usepackage{graphicx}
\usepackage{rotating} %Package added to allow the rotation of figures and chart on a page, {sidewaysfigure} command
\usepackage{tablefootnote} %Packaged added to allow footnotes in the tabular environment, use \tablefootnote command
\title{Freezing of a Weeks-Chandler-Anderson Fluid using classical Density Functional Theory}
\author{Kirstie L. Finster}
\degree{Master of Science}
\doctype{Thesis}
\department{Physics}
\depttype{Department}
\depthead{Head}
\major{Physics}
\advisor{David Roundy}
\submitdate{December 10, 2020}
\commencementyear{2021}

@@

\date{December 2020}

\abstract{
  We use Classical Density Functional Theory (cDFT) to predict the 
  freezing of a Weeks-Chandler-Anderson (WCA) fluid. A WCA fluid 
  consists of soft spheres that exhibit the repulsive portion of the 
  Lennard-Jones potential widely used to model the interaction of two 
  neutral atoms. Our cDFT functional is constructed using Soft Fundamental 
  Theory (SFMT) with Tarazona's tensor expression for $\Phi_3$, and Schmidt's
  erf model in which we have incorporated temperature dependent
  parameters to fit the results to those predicted by Monte-Carlo simulations.
  Our functional was successful in predicting the freezing of the WCA fluid 
  with a fairly good match to Monte-Carlo simulations. This result suggests 
  that SFMT may provide a strong foundation for future density functionals 
  that will model real fluids.
}

@@

Soft Fundamental Measure Theory (SFMT) will be used to form the 
functional for the excess Helmholtz free energy. 
SFMT is a method that makes the task of 
forming an expression for the excess Helmholtz free energy of a particular 
model fluid much easier since the interaction that each sphere, or atom, 
has with every other sphere does not have to be taken explicitly into account.
Once the functional for the Helmholtz free energy is formed, then 
according to cDFT, the equilibrium
Helmholtz free energy of the system can be found for a particular 
temperature and density by varying the number 
density profile $n(\vec{r})$ until the free energy is minimized. 
Phase diagrams can then be constructed using Maxwell's Double Tangent Construction. 
Results will be compared to those from Monte-Carlo simulations.

@@

The FMT equations developed so far have been for a Hard-sphere fluid.
All of the weight functions, the relations between them, and the equation 
for the derivative of the mayer function, can be expressed in terms of 
$w_{2}(r)$ which, for a Hard-sphere fluid, consists of a delta function.
The weight function $w_{2}(r)$ is in turn related to $w_{3}(r)$ through 
Equation~\ref{w2_w3_relation}.
Soft Fundamental Measure Theory extends FMT discussed so far to fluids 
with soft potentials by keeping the relationships between the weight 
functions (given by Equations~\ref{eq:w0},~\ref{eq:w1},~\ref{eq:w_v1},
~\ref{eq:w_v2},~\ref{w2_w3_relation},~\ref{eq:tensorweight}) and the 
equation for the derivative of the mayer function 
(given by Equation~\ref{mayerder=w2convolution}) the same, 
but replacing $w_{3}(r)$ with a weight function 
appropriate for a soft sphere potential~\cite{schmidt1999density, schmidt2000fluid}. 
This is what we do in forming a functional for the Weeks-Chandler-Anderson 
fluid as detailed in the next chapter on Methods.

@@

Our goal is to generate temperature-density and pressure-temperature 
phase diagrams for a Weeks-Chandler-Anderson model fluid. 
A temperature-density phase diagram for a fluid that does not
have attraction between the particles that make up the fluid will show
two distinct regions - a "liquid" region at lower densities, and a solid 
region at higher densities. 
A phase transition between a liquid and a solid is accompanied by an 
abrupt change in the number density, and so there is a gap between the 
purely liquid and the purely solid regions. Within the gap, or coexistence
region, is found mixtures of various amounts of liquid and solid.  
Within the coexistence region, the liquid and solid have number densities 
that are fixed at the values they had at the phase transition. To construct
the phase diagram, these liquid and solid phase transition number densities
must be found. Maxwell's double tangent construction can be used to find 
these number densities given plots of the liquid and the solid Helmholtz 
free energies verses the inverse density at a given temperature. 
To construct these plots we use classical Density Functional Theory 
to find the Helmholtz free energies of the liquid and the solid. 
A functional for the ideal portion of the
Helmholtz free energy is given by Equation~\ref{eq:Fideal-n(r)}, and a 
functional for 
the excess Helmholtz free energy, which depends on the interatomic potential
between particles, will be developed in the next section using SFMT. 
Once a functional is constructed, the Helmholtz free energy can be found - 
but not directly. The equilibrium number density profile must first be 
identified. According to cDFT, the equilibrium number density profile, 
and its corresponding equilibrium Helmholtz free energy can be found 
by trying many different number densities, and identifying the one that 
minimizes the Helmholtz free energy. In this section, we look at how to 
generate various number density profiles that can be used to implement cDFT.

It is known that a fluid with a soft potential like the WCA fluid will 
form a Face Centered Cubic (FCC) crystal in its solid state~\cite{Hansen}. 
Although the entire 
crystal lattice can be constructed with FCC cells, 
FCC cells are not the smallest cells with which the lattice can be built. 
The shaded parallelepiped shown within the FCC cell in Figure \ref{fig:primitivecell} 
is the smallest structure that, when repeated, re-creates the entire crystal 
lattice. This parallelepiped primitive cell takes up one-fourth the volume 
of an FCC cell, and contains the equivalent of one atom.
It includes portions of eight atoms each located on a lattice site conveniently 
reachable by moving along one of three different lattice vectors. 
The three lattice vectors will be useful in 
creating the lattice structure computationally, and will also make the 
program run faster.

  \begin{figure}[h!]
    \centering
    \includegraphics[height=3.5cm]{PrimitiveCellLightBlue.png}
    \caption{The Primitive Crystal Cell is shown by the shaded parallelepiped}
    \label{fig:primitivecell}
  \end{figure}

An atom will not realistically reside exactly at a lattice point. 
Statistically there will typically be a normal, or Gaussian distribution 
of positions about the lattice point. 
Of course, atoms are always in motion, so when talking about "the location" 
of an atom, we are talking about a time averaged position, so that it 
does not change with time.
Figure \ref{fig:Ensemble_Gaus} shows examples 
of four possible positions an atom (shown in blue) might have in a crystal 
cell that has one lattice site (represented by a black dot the center of 
the cell). The atom is slightly off the lattice site in each example, and 
a sampling of many possibilities - more than just the four given 
as examples - would create an average profile that looks like a Gaussian
distribution, as shown on the right.
A plot of a Gaussian distribution in 2-dimensions is shown in 
Figure~\ref{fig:Gaus_plot} where the width of the Gaussian is described 
by $\sigma$ (not to be confused with the $\sigma$ related to the 
Lennard-Jones model).

3-dimensional Gaussian distributions centered at lattice points can be 
used to represent a number density profile $n(\vec{r})$  which describes 
how the atoms are distributed on average over a collection of cells.

 \begin{figure}[h!]
    \centering
    \includegraphics[height=2.5cm]{Ensemble_Gaussian.png}
    \caption{Four examples of possible positions an atom (blue sphere) 
    might have within a cell that has one lattice site (black dot) at 
    the center are shown on the left. A large number of such 
    possibilities would give rise to a Gaussian distribution, 
    as shown on the right.}
    \label{fig:Ensemble_Gaus}
  \end{figure}

@@

There can be many different number density profiles for a given number
density. The number density is the number of atoms per volume, whereas
the number density profile describes how much "atom" will be found at various 
points in the cell. 
It can be thought of as an average over a large collection of actual cells
of various possible spatial distributions of the atoms within a cell, or
more precisely, as an ensemble average over all possible microstates.
\begin{align}
    n(\vec r)=~<\sum_{i=1}^N\delta(\vec r - \vec r_i)>~
\end{align} 
By varying the width of the Gaussian distributions, a number of different 
number density profiles can be generated for a given number density.
It is important to note that the number density does not change when 
introducing a number density profile.  
There are still the same number of atoms over some fixed volume. 
Taking the average of the number density profile over some volume 
gives the number density. 
\begin{align}
    n=<n(\vec{r})>=\frac{1}{V}\int{n(\vec{r})}{d\vec{r}}=\frac{N}{V}
\end{align}

Besides changing the width of the Gaussians, there is another way to 
generate different number density profiles for a given number density.
It may be that not all lattice sites are occupied. Vacancies can be taken 
into account while keeping the overall number density the same by 
making the cells smaller. A large collection of smaller cells creates a 
new number density profile shown in Figure~\ref{fig:Ensemble_Smallcells} 
with eight smaller Gaussians each of which is now one-eighth the height. 
Instead of thinking of an atom in one of the eight small cells, the same 
Gaussian distribution can be obtained from thinking of one-eighth of an 
atom in each small cell as shown in Figure~\ref{fig:SameStatPic}. 
This makes it convenient to describe the new 
number density in terms of the fraction of vacancies, $f_v$: 
\begin{displaymath}{ n = (1-f_v){n_{no.vacancies}}}\end{displaymath}

 \begin{figure}[h!]
    \centering
    \includegraphics[height=2.5cm]{Ensemble_Smallcells.png}
    \caption{Four examples of possible positions an atom (blue sphere) might 
    have within a fixed volume containing eight cells are shown on the left. 
    Each cell has one lattice site at the center represented by a black dot. 
    A large number of such possibilities would give rise to the Gaussian 
    distributions shown on the right.}
    \label{fig:Ensemble_Smallcells}
  \end{figure}

@@

By varying the width of the Gaussians and the fraction of vacancies, 
various number density profiles can be generated for a given combination 
of temperature and number density, as shown in Figure~\ref{fig:Ensemble_vary}.
A homogeneous number density profile is used for a liquid is shown in 
Figure \ref{fig:homogen_denisty}.
Once again, it is important to note that the number density is the same 
for all these number density profiles $n(\vec{r})$. 
\begin{figure}[h!]
    \centering
    \includegraphics[height=8cm]{VaryWidthandVacancies.png}
    \caption{Several Gaussian Distribution Number Density Profiles all 
    with the same number density of one atom over the fixed volume.}
    \label{fig:Ensemble_vary}
  \end{figure}  
 \begin{figure}[h!]
    \centering
    \includegraphics[height=4cm]{figs/homogeneous_bold-box.pdf}
    \caption{A homogeneous number density profile $n(\vec{r})$.}
    \label{fig:homogen_denisty}
  \end{figure}   
The Helmholtz free energy $F[n(\vec{r})]$ is calculated for each number 
density profile, and also for the homogeneous number density profile 
characteristic of a liquid. Both the ideal and excess free energies are 
found, and added together to get the total Helmholtz free energy.
According to classical Density Functional Theory, of all the Helmholtz 
free energies calculated is the minimum Helmholtz free energy, and its 
corresponding number density profile that are the actual equilibrium values 
for that temperature and number density.
The process is repeated to find the free energies at other combinations 
of temperatures and number densities.

Once the Helmholtz free energy for both the crystal and liquid states of 
the fluid are found for a given temperature and density, they can be 
used to construct phase diagrams.
Outside the region of coexistence which exists between the liquid and 
crystal states, the fluid will be completely in a crystalline state if the 
crystal free energy is lower than the liquid free energy, and it will be 
completely in a liquid state if the liquid free energy is lower than 
the crystal free energy. Maxwell's Double Tangent Construction is used 
to determine the actual number densities of the crystal and liquid states
at the boundaries of the coexistence region for a given temperature, 
and construct phase diagrams as described in Section 2.2.

@@

Phase diagrams for the WCA fluid are shown in Figures~\ref{fig:Phase_Diagram_of_T_vs_n} 
and \ref{fig:Phase_Diagram_P_vs_T}. In all cases $\sigma=1$ and $\epsilon=1$. 
In Figure~\ref{fig:Phase_Diagram_of_T_vs_n} the reduced temperature 
$T^*=\frac{kT}{\epsilon}$ is plotted against the reduced number density 
$n^*=n\sigma^3$ for temperatures of kT = 0.5 to 3.0. 
It is seen that in general, a liquid exists at lower number densities, 
and a solid exists at higher number densities. 
Phase transitions between the liquid and solid phases will occur when the liquid 
and solid are at the same temperature, pressure and chemical potential. 
Looking at one temperature, it is seen that as the density of the 
liquid increases, a maximum number density is reached for the liquid. 
This is the liquid phase transition number density, and 
marks the beginning of a coexistence region where various amounts 
of liquid and solid can exist together in equilibrium.
Moving on toward higher number densities, the end of the coexistence 
region is marked by the minimum number density that the solid can have.
This is the solid phase transition number density. At a phase transition 
there is an abrupt change in the number density as seen here where the
density changes suddenly, and abruptly from the liquid 
phase transition number density to the solid phase transition number density.
Within the coexistence region, the number density of the liquid is fixed 
at the liquid phase transition number density, and the number density of 
the solid is fixed at the solid phase transition number density, while
varying amounts of liquid and solid combine to give the infinite number 
of possible densities seen within the coexistence region.

\begin{figure}
  \centering
  \includegraphics[height=9cm]{figs/Phase_Diagram_of_T_vs_n}
  \caption{Phase Diagram of the reduced temperature verses the reduced 
  density for the WCA fluid. The red region is liquid, and the blue region is 
  solid. The gray area that separates the liquid and solid regions 
  is the region of coexistence where some liquid and solid exist together 
  at varying proportions. White lines show the boundaries of the coexistence 
  region generated from Monte-Carlo of the WCA fluid.}
  \label{fig:Phase_Diagram_of_T_vs_n}
\end{figure}

Looking at the coexistence region as a whole for all temperatures, 
it is seen to extend over a narrow range of number densities which 
gradually shifts towards higher densities as the temperature increases. 
This is expected since at higher temperatures the fluid would have to 
be under greater compression before forming a solid. 
White lines superimposed on the phase diagram show the boundaries of the 
coexistence region as predicted by Monte-Carlo simulations~\cite{May} 
of the WCA fluid. 
Overall, the general shape, and position of the two coexistence regions 
correspond fairly closely, but the Monte-Carlo region is a little wider
and shifted a bit toward smaller densities. For example, at a temperature 
of kT=1, the cDFT liquid transition reduced number density is 0.96, 
and the solid transition reduced number density is 1.02, giving a spread 
of 0.06, whereas the Monte-Carlo results give a liquid transition 
reduced number density of 0.92, and a solid transition reduced number 
density of 1.00, giving a spread of 0.08. 
%The solid phase transition number densities predicted by each method 
%are very nearly the same along all temperatures, but the liquid phase 
%transition number densities predicted by cDFT are notably higher than 
%those predicted by the Monte-Carlo simulations, making the width of the 
%Monte-Carlo region about one-third larger for most temperatures above KT=1. 
%Figure~\ref{fig:Phase_Diagram_of_T_vs_n_AS} shows our results for a wider 
%range of reduced temperatures from kT=0.5 to 38. The white lines show the results
%of molecular simulations of the WCA fluid made by Ahmed and Sadus~\cite{ahmedsadus}.
%As can be seen in the figure, the region of coexistence in our simulations 
%falls within the region of coexistance predicted by Ahmed and Sadus at almost 
%all temperatures from approximately kT=0.5 to 28. At the low temperature of kT=0.5,
%the width of the coexistance regions are nearly the same, while moving to 
%higher temperatures,the width of our coexistance region ranges from about 
%one-half to one-forth the size. Although, our coexistence region is only 
%about one-forth the size at kT=28, it is centered within the coexistence 
%region predicted by Ahmed and Sadus. 
%
%
%
%\begin{figure}
  %\centering
  %%\includegraphics[width=\columnwidth]{figs/Phase_Diagram_of_T_vs_n_AS}
  %%\includegraphics[height=10cm]{figs/Phase_Diagram_of_T_vs_n_AS}
  %%\caption{Phase Diagram of the reduced temperature verses the reduced density for the WCA fuild. 
  %%White lines show the liquid and solid transition densities at each temperature from 
  %%simulations of a WCA fluid by Ahmed and Sadus~\cite{ahmedsadus}. The gray region is the region of coexistence.}
  %\label{fig:Phase_Diagram_of_T_vs_n_AS}
%\end{figure}

\begin{figure}
  \centering
  \includegraphics[height=9cm]{figs/Phase_Diagram_of_P_vs_T}
  \caption{Phase Diagram of the reduced pressure verses the reduced 
  temperature. The blue region is solid, and the red region is liquid.
  The white line marks the solid-liquid boundary as predicted by 
  Monte-Carlo simulations.}
  \label{fig:Phase_Diagram_P_vs_T}
\end{figure}

Figure~\ref{fig:Phase_Diagram_P_vs_T} shows the reduced pressure 
$p^*=\frac{P\sigma^3}{\epsilon}$ plotted against the reduced temperature 
for temperatures of kT = 0.5 to 3.0. 
The black line marks the phase boundary between the solid and liquid regions, 
and indicates the various temperature-pressure combinations at which a 
phase transition occurs, and where various amounts of solid and liquid 
can coexist. 
It is seen that the higher the temperature, the higher the pressure must be
to undergo a phase transition from a liquid to a solid, as would be expected. 
The slope of the phase boundary $\frac{dP}{dT}$ at some point (P,T) is given by
\begin{align} 
    \frac{dP}{dT}=\frac{S_l-S_s}{V_l-V_s}
\end{align}
Since the entropy and the volume of the liquid is greater than the entropy
and the volume of the solid, the slope should be positive.
Comparing the slope of the phase boundary predicted by cDFT to that
from Monte-Carlo simulations shown by the white line, it is seen that
the cDFT transition temperatures are consistently higher, and the 
discrepancy grows as the temperature increases. For example, at a 
temperature of kT=1, the phase transition reduced pressure predicted by 
cDFT is 15.24, whereas the Monte-Carlo simulation predicted a value of 9.99,
and at a temperature of kT=3, cDFT gives a value of 77.27 whereas the 
Monte-Carlo simulation predicts 43.80.
This may indicate that either the difference between the liquid 
and solid phase transition entropies predicted by cDFT is too high, or 
the difference in the volumes is too low. 
%
%Figure~\ref{fig:Phase_Diagram_P_vs_T_AS} shows our results for a larger 
%range of reduced temperatures from kT=0.5 to 38, compated with the results
%from Ahmed and Sadus's molecular simulations. Here again, our transition 
%pressures and temperatures are consistently higher.
%
%
\begin{figure}
  \centering
  %\includegraphics[width=\columnwidth]{figs/Phase_Diagram_of_P_vs_T_AS}
  %\includegraphics[height=10cm]{figs/Phase_Diagram_of_P_vs_T_AS}
  %\caption{Phase Diagram of the reduced pressure verses the reduced temperature. 
  %The white line marks the boundary from simulations by Ahmed and Sadus~\cite{ahmedsadus}.}
  \label{fig:Phase_Diagram_P_vs_T_AS}
\end{figure}
%
%Figure~\ref{fig:p-vs-T_at_fixed_density} shows individual plots of the 
%reduced pressure versus the reduced temperature at fixed reduced densities 
%from $n^*$ = 0.7 to 1.2. As can be seen by comparison with Figure 
%\ref{fig:Phase_Diagram_P_vs_T_AS}, the plotted lines display kinks when 
%the WCA fluid transistions into a solid. These downward bends align to 
%form the curved interface separating the blue and red areas in 
%Figures~\ref{fig:Phase_Diagram_P_vs_T} and \ref{fig:Phase_Diagram_P_vs_T_AS}.
%
\begin{figure}
  \centering
  %\includegraphics[width=\columnwidth]{figs/p-vs-T_at_fixed_n_noMC}
  %\caption{Plot of the reduced pressure verses the reduced temperature. 
  %The lines bending downward indicate freezing.}
  \label{fig:p-vs-T_at_fixed_density}
\end{figure}
